---
title: "Practicegithub"
author: "Jacob Mehlhoff"
date: "March 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First the libraries must be loaded.

```{r, include = FALSE}
library(readxl) 
library(tidyverse) 
```

The finches data must be read

```{r}
finches <- read_excel("finches_data.xlsx")
```

Look at the data.

```{r}
glimpse(finches)
```

Create a histogram of beak length grouped by survival.

```{r}
ggplot(
  data = finches,                     
  mapping = aes(x = beak_length,      
                fill = outcome)       
) +
  geom_histogram(bins = 14) +        
  facet_wrap(~ outcome, ncol = 1) +  
  guides(fill = FALSE) +              
  labs(
    title = "Figure 1.",             
    x = "Beak Length (mm)",           
    y = "Number of Birds"             
  )
```

Summarize the data set by outcome.

```{r}
beak_length_grouped_summary <- 
  finches %>% 
  group_by(outcome) %>% 
  summarize(mean = mean(beak_length),
            sd = sd(beak_length),
            n = n()) %>% 
  mutate(sem = sd / sqrt(n),
         upper = mean + 1.96 * sem,
         lower = mean - 1.96 * sem)
```

Create a bar chart.

```{r}
ggplot(
  data = beak_length_grouped_summary,   # dont use the original finches dataset
  mapping = aes(x = outcome,            # survival on the x axis
                y = mean,               # mean beak length on the y axis
                fill = outcome)         # make died/survived different colors
) +
  geom_col() +                          # add columns
  geom_errorbar(                        # add error bars
    mapping = aes(ymin = lower,         #   lower 95% confidence limit
                  ymax = upper),        #   upper 95% confidence limit
    width = .3                          #   width of horizontal part of bars
  ) +
  guides(fill = FALSE) +                # don't show a legend for fll color
  labs(
    title = "Figure 2.",                # title
    x = "Survival Outcome",             # x-axis label
    y = "Beak Length (mm)"              # y-axis label
  )
```

Get a vector of beak lengths for birds that died.

```{r}
beak_length_died <-
  finches %>%                     # start with finches dataset
  filter(outcome == "died") %>%   # only include rows w/ outcome=died
  pull(beak_length)               # extract the beak_length column

```

Print the new object in the console... it is a vector
```{r}
beak_length_died
```

Get a vector of beak lengths for birds that survived.

```{r}
beak_length_survived <-
  finches %>% 
  filter(outcome == "survived") %>% 
  pull(beak_length)
```

Print the results in the console.

```{r}
beak_length_survived
```

Perform a two-sample t-test assuming unequal variances.

```{r}
t.test(beak_length_died, beak_length_survived)
```

